---
title: "Space Time"
output: html_notebook
---

# Libraries
```{r}
library(sf)
library(ggplot2)
```

# WD
```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data")

```


# Generating centroids
```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data")

#Import my shapefile
ec3 <-st_read(dsn="C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data",layer="nxprovincias")
ec3 <- subset(ec3, DPA_DESPRO != "GALAPAGOS")

# ec3<-rgdal::readOGR(dsn="~/Dropbox/0.USP/10.2020 II sem/FLI/case-control/bayesian/",layer="nxparroquias")
ec3 <- st_transform(ec3, crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
trueCentroids <- st_centroid(ec3)
trueCentroids_df <- as.data.frame(st_coordinates(trueCentroids))

#Add centroids to the shapefile
ec3$lon <- trueCentroids_df$X
ec3$lat <- trueCentroids_df$Y

ec <- data.frame(ec3)


# Checking locations
ggplot()+
  geom_sf(data = ec3, fill= "lightblue", color = "black") +
           geom_point(data=ec, aes(x=lon, y=lat),
                      size=1, shape=21, fill="orange", color="black")

ec <- st_drop_geometry(ec3)
write.csv(ec, file = "nxprovincias.csv", row.names = FALSE)

str(ec)


# Parroquias
library(sf)

# Read the shapefile
ec3 <- st_read(dsn = "C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data", layer = "nxparroquias")

# Subset to exclude "GALAPAGOS"
ec3 <- subset(ec3, DPA_DESPRO != "GALAPAGOS")

# Ensure geometries are valid
if (!all(st_is_valid(ec3))) {
  ec3 <- st_make_valid(ec3)  # Fix invalid geometries
}

# Transform to WGS84
ec3 <- st_transform(ec3, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

# Calculate centroids
trueCentroids <- st_centroid(ec3)

# Handle the st_centroid warning
# Disaggregate multipart geometries if needed
if (any(st_geometry_type(ec3) == "GEOMETRYCOLLECTION")) {
  ec3 <- st_collection_extract(ec3, "POLYGON")
}

# Extract centroid coordinates and add to the shapefile
trueCentroids_df <- as.data.frame(st_coordinates(trueCentroids))
ec3$lon <- trueCentroids_df$X
ec3$lat <- trueCentroids_df$Y

# Convert the sf object to a data frame
ec <- data.frame(ec3)

# Print summary
print(head(ec))


# Checking locations
ggplot()+
  geom_sf(data = ec3, fill= "lightblue", color = "black") +
  geom_point(data=ec, aes(x=lon, y=lat),
             size=1, shape=21, fill="orange", color="black")

ec <- st_drop_geometry(ec3)
write.csv(ec, file = "nxparroquias.csv", row.names = FALSE)

str(ec)



```

# geo import centroids and id
```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data")
geo <- read.csv(file = "nxparroquias.csv", colClasses = "character")
length(unique(geo$DPA_PARROQ))
```

# pop population file
# Reading censo avicola

```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data")
c <- readxl::read_excel("Base Datos_Censo_Avicola-2015-10-21-1.xlsx")
```

# are all unique ?
```{r}
#There are 2206 unique num_cuestionario it was supposed to be unique but not there are 84 duplicated, maybe errors, maybe something else. 
length(unique(c$NUM_CUESTIONARIO))
table(duplicated(c$NUM_CUESTIONARIO))

#  Creating a new unique number to avoid the 84 duplicated only 1 found that it was different as well
table(duplicated(paste(c$IDUBICACION, c$NUM_CUESTIONARIO)))

# Creating a new id for grouping purpouses and avoid complication
c$id <- seq(1, nrow(c)) 
```

# grouping censo avicola
```{r}
pop <- c %>% 
  group_by(id=IDUBICACION) %>% 
  summarise(granjas=n(),
            poblacion=sum(as.numeric(NUM_POLLOS), 
                          as.numeric(REP_LIVIANAS),
                          as.numeric(REP_PESADAS),
                          as.numeric(PONEDORAS),
                          as.numeric(PAVOS)))
length(unique(pop$id))
table(pop$granjas)
table(pop$poblacion)

sum(pop$granjas[pop$id_check == "TRUE"]) #there are 2031 farms
sum(pop$granjas[pop$id_check != "TRUE"]) #there are 259 farms

table(pop$id %in% geo$DPA_PARROQ)
pop$id_check <- pop$id %in% geo$DPA_PARROQ

# Filtering the ones that match in the geo file
pop <- pop %>% 
  filter(id_check == TRUE)

pop <- rbind(pop, c("020155", "1", "0", "NA"))
pop <- rbind(pop, c("020159", "1", "0", "NA"))
pop <- rbind(pop, c("020153", "2", "0", "NA"))
pop <- rbind(pop, c("020157", "1", "0", "NA"))
pop <- rbind(pop, c("050157", "1", "0", "NA"))
pop <- rbind(pop, c("050153", "1", "0", "NA"))
pop <- rbind(pop, c("170250", "1", "0", "NA"))
pop <- rbind(pop, c("180165", "1", "0", "NA"))
pop <- rbind(pop, c("180853", "1", "0", "NA"))

pop <- data.frame(pop)

```
New 

# cases cases file
```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data")
casos <- readxl::read_excel("casos parroquias.xlsx")
str(casos)
casos$codigo_parroquia <- substr(casos$codigo_parroquia, 1,6)

ggplot(casos, aes(x=lon_wahis, y=lat_wahis))+
      geom_point()

plotly::ggplotly(ggplot(casos, aes(x=lon_wahis,y=lat_wahis))+
      geom_point())


```

```{r}
cases <- casos %>% 
  group_by(id=codigo_parroquia, date=`Fecha_1er_ave enferma`) %>% 
  summarise(cases_granjas=n(),
            cases_aves=sum(as.numeric(`aves afectadas`)))

cases <- data.frame(cases)
```

#checking ID in geo
```{r}
table(cases$id %in% geo$DPA_PARROQ)

cases$check_in_geo <- cases$id %in% geo$DPA_PARROQ

cases <- cases %>%
  filter(cases$check_in_geo == TRUE)

cases <- data.frame(cases)

cases$check_in_pop <- cases$id %in% pop$id
table(cases$check_in_pop)
```


# writing csv for sat scan
```{r}
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data/satscan/")

write.csv(pop, file = "pop.csv")
write.csv(cases, file = "cases.csv")
write.csv(geo, file = "geo.csv")

```

# You are running SaTScan v10.1.3.

# SaTScan is free, available for download from https://www.satscan.org/.
# It may be used free of charge as long as proper citations are given
# to both the SaTScan software and the underlying statistical methodology.
# 
# Reading the coordinates file
# Reading the population file
# Reading the case file
# Calculating the expected number of cases
# Constructing the circles
# Finding the most likely clusters
# .... this step will take approximately 9 minutes.
# .... this step will take approximately 8 minutes more.
# .... this step will take approximately 5 minutes more.
# .... this step will take approximately 3 minutes more.
# Checking the Overlapping Nature of Clusters
# SaTScan log likelihood ratio for the most likely cluster:   51.56
# Doing the Monte Carlo replications
# SaTScan log likelihood ratio for #1 of 999 replications:   10.81
# SaTScan log likelihood ratio for #2 of 999 replications:    8.33
# SaTScan log likelihood ratio for #3 of 999 replications:   10.08
# SaTScan log likelihood ratio for #4 of 999 replications:   10.47
# .... this step will approximately take at most 22 hours 10 minutes.
# SaTScan log likelihood ratio for #5 of 999 replications:   13.13
# SaTScan log likelihood ratio for #6 of 999 replications:   15.95
# SaTScan log likelihood ratio for #7 of 999 replications:   12.28
# SaTScan log likelihood ratio for #8 of 999 replications:   11.90
# SaTScan log likelihood ratio for #9 of 999 replications:    8.78
# SaTScan log likelihood ratio for #10 of 999 replications:   10.42
# SaTScan log likelihood ratio for #11 of 999 replications:   13.52
# SaTScan log likelihood ratio for #12 of 999 replications:   11.96
# SaTScan log likelihood ratio for #13 of 999 replications:   12.93
# SaTScan log likelihood ratio for #14 of 999 replications:   10.12
# SaTScan log likelihood ratio for #15 of 999 replications:   11.25
# SaTScan log likelihood ratio for #16 of 999 replications:   10.01
# SaTScan log likelihood ratio for #17 of 999 replications:    8.58
# SaTScan log likelihood ratio for #18 of 999 replications:    8.87
# SaTScan log likelihood ratio for #19 of 999 replications:   17.39
# SaTScan log likelihood ratio for #20 of 999 replications:   12.19
# SaTScan log likelihood ratio for #21 of 999 replications:   10.80
# SaTScan log likelihood ratio for #22 of 999 replications:    9.67
# SaTScan log likelihood ratio for #23 of 999 replications:    8.37
# SaTScan log likelihood ratio for #24 of 999 replications:   12.81
# SaTScan log likelihood ratio for #25 of 999 replications:    9.53
# SaTScan log likelihood ratio for #26 of 999 replications:   11.24
# SaTScan log likelihood ratio for #27 of 999 replications:    9.90
# SaTScan log likelihood ratio for #28 of 999 replications:   11.81
# SaTScan log likelihood ratio for #29 of 999 replications:   15.68
# SaTScan log likelihood ratio for #30 of 999 replications:   13.02
# SaTScan log likelihood ratio for #31 of 999 replications:   11.02
# SaTScan log likelihood ratio for #32 of 999 replications:    9.77
# SaTScan log likelihood ratio for #33 of 999 replications:    8.26
# SaTScan log likelihood ratio for #34 of 999 replications:   13.67
# SaTScan log likelihood ratio for #35 of 999 replications:    9.82
# SaTScan log likelihood ratio for #36 of 999 replications:    9.04
# SaTScan log likelihood ratio for #37 of 999 replications:    9.01
# SaTScan log likelihood ratio for #38 of 999 replications:   15.79
# SaTScan log likelihood ratio for #39 of 999 replications:   11.37
# SaTScan log likelihood ratio for #40 of 999 replications:   11.15
# SaTScan log likelihood ratio for #41 of 999 replications:    8.87
# SaTScan log likelihood ratio for #42 of 999 replications:   10.32
# SaTScan log likelihood ratio for #43 of 999 replications:   11.56
# SaTScan log likelihood ratio for #44 of 999 replications:    9.30
# SaTScan log likelihood ratio for #45 of 999 replications:    9.92
# SaTScan log likelihood ratio for #46 of 999 replications:    9.74
# SaTScan log likelihood ratio for #47 of 999 replications:   11.63
# SaTScan log likelihood ratio for #48 of 999 replications:   12.92
# Cancelling job, please wait while execution is interrupted ...

```{r}

```


Plotting the clusters
```{r}
library(ggmap)
library(ggspatial)
library(sf)
library(ggforce)

# API registration 
register_stadiamaps(key = "76b5ee9f-672a-4475-95ac-c3e2cbfbb72a")

# Ecuador area  with cluster
cl <-read_sf("C:/Users/alfredo.acosta/OneDrive - SVA/ASF_Uganda/R/Spatiotemporal/61retrospec_min5locations/results.col.shp")

plot(cl)

# Use the lu map already created
ec <- ggmap::get_stadiamap(bbox = c(left = -81.5,
                                    right = -75,
                                    bottom = -5.1, 
                                    top = 1.5),
                                    zoom = 7, #zoom 14 was ok
                                    maptype = "stamen_terrain")
ggmap(ec)

ggmap(ec) + 
  geom_point(data=casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis)), 
             color = 'red',
             size= 0.5)

```
# Read Shapefile 
```{r}
# cluster
cluster_ec <-read_sf("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data/satscan/results_space3_granjas.gis.shp")

plot(cluster_ec)
```
```{r}
cluster <- ggmap(ec) +
    ggforce::stat_circle(aes(x0 = cluster_ec$LOC_LONG[1], 
                         y0 = cluster_ec$LOC_LAT[1], 
                         r = 14.04/111), colour="purple", alpha=0.01, fill= "red") + 
    ggforce::stat_circle(aes(x0 = cluster_ec$LOC_LONG[10], 
                         y0 = cluster_ec$LOC_LAT[10], 
                         r = 52.52/111), colour="purple", alpha=0.01, fill= "purple") +
    ggforce::stat_circle(aes(x0 = cluster_ec$LOC_LONG[626], 
                         y0 = cluster_ec$LOC_LAT[626], 
                         r = 18.83/111), colour="purple", alpha=0.01, fill= "green") +
    ggforce::stat_circle(aes(x0 = cluster_ec$LOC_LONG[660], 
                         y0 = cluster_ec$LOC_LAT[660], 
                         r = 22.43/111), colour="purple", alpha=0.01, fill= "yellow") +
    geom_point(data=casos, aes(x=as.numeric(lon_wahis), 
                               y=as.numeric(lat_wahis)), 
             color = 'red',
             size= 0.5) +
  labs(x=NULL, y=NULL, fill="Space-time clusters") +
  scale_fill_manual(values=c("1. RR:13.8 (t=37 days)"="red",
                             "2. RR:18.6 (t= 5 days)"="purple",
                             "2. RR:18.6 (t= 5 days)"="green",
                             "2. RR:18.6 (t= 5 days)"="yellow"
                             )) +
  theme(legend.position = "right")
```


```{r}
cluster_fig <- cluster +
  annotation_scale(location= "br") +
  coord_sf(crs = 4326)



# Export map
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data/satscan/")

tiff(filename = "Fig-space-3.tiff", width=13, height=10,
     units="cm", res=1600,compression = "lzw", pointsize = 12)
cluster_fig
dev.off()


```




# Kernel heat map
## general
```{r}
# Set sys locale to avoid swedish months
Sys.setlocale("LC_TIME", "C")

# Now, use the data in your plot
ec_kernel <- 
  ggmap(ec) +
  stat_density_2d(data = casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis), fill = ..level..),
                  geom = "polygon", alpha = 0.4) +  # Heatmap using density
  scale_fill_viridis_c() +  # Color scale for heatmap
    geom_point(data=casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis)), 
             color = 'red',
             size= 0.5,
             alpha=0.5)

  ec_kernel

# limiting the view    
  # + 
  # xlim(31.92, 32.1) +  # Adjust x-axis limits
  # ylim(2.44, 2.6) +    # Adjust y-axis limits
  # theme_void() +         # Remove background and axis

ec_k <- ec_kernel +
  annotation_scale(location= "br") +
  coord_sf(crs = 4326)
ec_k

# Export map
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data/kernel/")

tiff(filename = "Fig-map_kernel.tiff", width=13, height=10,
     units="cm", res=1600,compression = "lzw", pointsize = 12)
ec_k
dev.off()

```
## Montly seasonality
```{r}
Sys.setlocale("LC_TIME", "C")

# generating montly variable to see any kind of temporal aggregation
casos <- casos %>%
  mutate(month_mod=month(`Fecha_1er_ave enferma`))

# geneate kernel by month
ec_kernel <- 
  ggmap(ec) +
  stat_density_2d(data = casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis), fill = ..level..),
                  geom = "polygon", alpha = 0.4) +  # Heatmap using density
  scale_fill_viridis_c() +  # Color scale for heatmap
    geom_point(data=casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis)), 
             color = 'red',
             size= 0.5) +
  facet_wrap(
    ~ month_mod, 
    nrow = 4, ncol = 4)

  + 
  xlim(31.92, 32.1) +  # Adjust x-axis limits
  ylim(2.44, 2.6) +    # Adjust y-axis limits
  theme_void() +         # Remove background and axis
  theme(legend.position = "none")

lmk <- lutuk_kernel +
  annotation_scale() +
  # annotate("text", label = "Lutuk", y = 2.56, x = 32.05, size = 3.5, colour = "black") +
  coord_sf(crs = 4326)
lmk
# Export map (off)
# setwd("C:/Users/alfredo.acosta/OneDrive - SVA/ASF_Uganda/Results/Figures/")
# 
# tiff(filename = "Fig-map6_kernel.tiff", width=13, height=10,
#      units="cm", res=1600,compression = "lzw", pointsize = 12)
# lmk
# dev.off()

```


Semester seasonality
```{r}
# Creating a semester variable
casos <- casos %>%
  mutate(
    semester = case_when(
      month(as.Date(`Fecha_1er_ave enferma`)) %in% 1:6 ~ paste(year(as.Date(`Fecha_1er_ave enferma`)), "1st Semester"),
      month(as.Date(`Fecha_1er_ave enferma`)) %in% 7:12 ~ paste(year(as.Date(`Fecha_1er_ave enferma`)), "2nd Semester")
    ),
    semester_order = case_when(
      month(as.Date(`Fecha_1er_ave enferma`)) %in% 1:6 ~ year(as.Date(`Fecha_1er_ave enferma`)) + 0.0,
      month(as.Date(`Fecha_1er_ave enferma`)) %in% 7:12 ~ year(as.Date(`Fecha_1er_ave enferma`)) + 0.5
    )
  ) %>%
  arrange(semester_order) %>% # Arrange data by temporal order
  mutate(semester = factor(semester, levels = unique(semester))) # Create an ordered factor

Sys.setlocale("LC_TIME", "C")

# geneate kernel by month
ec_kernel_sem <- 
  ggmap(ec) +
  stat_density_2d(data = casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis), fill = ..level..),
                  geom = "polygon", alpha = 0.4) +  # Heatmap using density
  scale_fill_viridis_c() +  # Color scale for heatmap
    geom_point(data=casos, aes(x=as.numeric(lon_wahis), y=as.numeric(lat_wahis)), 
             color = 'red',
             size= 0.1) +
  facet_wrap(
    ~ semester, 
    nrow = 2, ncol = 2)

  xlim(-80, -77) +  # Adjust x-axis limits
  ylim(0.8, -3.2)    # Adjust y-axis limits
  
ec_k_sem <- 
  ec_kernel_sem +
  annotation_scale(location= "br") +
  coord_sf(crs = 4326)
ec_k_sem

# Export map
setwd("C:/Users/alfredo.acosta/OneDrive - SVA/Papers/HPAI-Ecuador/data/kernel/")

tiff(filename = "Fig-map_kernel_sem.tiff", width=15, height=10,
     units="cm", res=3600,compression = "lzw", pointsize = 12)
ec_k_sem
dev.off()

```

